---
title: "Simple-guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simple-guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  out.width = '80%',
  fig.align = "center",
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MBSE.utils)
library(ggplot2)
library(ggnetwork)
```


## Traversing An XML File

```{r traversing}

# Load the sample XML file saved in this package:

test.xml.file.path <- system.file("extdata","Example1.xml", package = "MBSE.utils")
test.xml <- xml2::read_xml(test.xml.file.path)
xml.atts.list <- get.xml.attrs(test.xml)

xml.atts.list$xml_atts_sorted

```

There are `r xml.atts.list$number_of_attrs` attributes: `r xml.atts.list$all_attributes`

## Converting A File To Datum Triples

DatumTriples are the smallest unit of data.

```{r makingDatumTriples}
# Using test.xml again
test.xml.file.path <- system.file("extdata","Example1.xml", package = "MBSE.utils")
test.xml <- xml2::read_xml(test.xml.file.path)

 all.nodes <- xml2::xml_find_all(test.xml, "/*", flatten = TRUE) %>% 
   set.datum.entity() # Set datum entity adds missing IDs to nodes.

# In complex files you might end up with a list of nodes sone need to traverse
# the list. Here we have only one, but show the index for clarity.

full.triple.out <- all.nodes[1] %>%  make.datum.triple()

# Print a sample
full.triple.out %>% head(5) %>% pander::pandoc.table()

```

# Visualisation

Now for a couple of comments about practical choices made for visualisation. The first is the use of the `igraph` package which is, without doubt, the most complex, but complete package of network related functions, so the least forgiving to use. This package supports a very wide range of statistical functions for characterising networks. Later work might needs these, for example, centralisation measures might relate to importance in the real world. Statistical clustering may also represents real world groups. While this is speculative, the choice of `igraph` keeps the option open.

The second point is the convoluted path of using `ggplot2` for plotting the results rather than using the `igraph` built-in plot methods. This has been done to leverage the 'grammar of graphics' [@Wilkinson2010] implementation within `ggplot2` ecosystem of packages. This effectively allows the instructions to produce a style of plot to be wrapped in a function independent of data. This abstraction is easy to apply to suitable data and avoids the need for a detailed understanding of the base package plot elements. The less wordy summary is: `ggplot2` methods are easy repeat across different data. 

```{r include=TRUE, echo=TRUE, results='asis', eval=TRUE}

# Using test.xml again
test.xml.file.path <- system.file("extdata","Example1.xml", package = "MBSE.utils")
test.xml.path <- fs::path_dir(test.xml.file.path)


list.xml <- fs::dir_ls( path = test.xml.path,
                           type="file", recurse = TRUE,
                           glob = "*.xml") # only these file types
# Get one file
file_name_used <- list.xml[1] # Select file from list
the_file_name <- fs::path_file(file_name_used) # Strip the path
clean_xml <- xml2::read_xml(file_name_used)   %>% xml2::xml_ns_strip()  #

# Lazy print the filename
the_file_name
```

## Finding The Diagram Elements

In the following code we find all the diagram elements using **Xpath** syntax. The presumption here is that the file contains one or more diagrams as that is the focus of our interest, and also that each selected element is an entry point for a hierarchy of child tags that form the diagram. We could choose any arbitrary tag type to make our entry point, but some make more sense than others. A good way to look for valid names is to open the raw XML file with your preferred web browser.


```{r}
# This lists all the diagrams
diagram.elements.all <-   clean_xml  %>% xml2::xml_find_all("//diagram") 

# Lazy print the number of diagrams
length(diagram.elements.all)
```

## Labelling The File Elements

The custom function `make.datum.triple()` takes a single xml node as a starting point and recurses down through all the child elements transforming them into `datumTriples`. The simplicity of the 'Tidy Data' method comes into its own here as we simply pipe the functions `set.datum.entity`, `set.datum.parent`, and `make.datum.triple` to act upon `diagram.to.label` to make `all.triples`

```{r include=TRUE, echo=TRUE, results='asis', eval=TRUE}

diagram.to.label <- diagram.elements.all[1] 
all.triples <- diagram.to.label %>% 
   set.datum.entity() %>% # Set the datumEntity
  set.datum.parent() %>%  # Set the parent of the node
  make.datum.triple() %>%  # Turn in to triples
  get.tag.value.pairs() # Expand Tag Value Pairs
# Lazy print diagram.to.label
diagram.to.label

```

## Make Into Rectangular Format And Plot

The simple visualisation of the XML structure below matches the form expected by manual inspection. 

```{r manualCodeBlock, include=TRUE, echo=TRUE, results='markup', eval=TRUE, fig.cap= "Plot generated by manual code block"}

wide.data.ss <- all.triples %>% triple.as.wide()

# Make the vertices (Nodes)

network.data.vertices <- wide.data.ss %>% dplyr::select(datumEntity,tag.type)


# Add a couple of columns for edges

wide.data.ss$from <- wide.data.ss$datumParent
wide.data.ss$to <- wide.data.ss$datumEntity

# The entity at the top of the tree does not have a parent, so is marked base
# Drop the base link. This would work if we combined multiple diagrams
network.data.edges<- wide.data.ss %>% dplyr::select(from,to) %>%   #filter(from != "Base" )
dplyr::filter(!stringr::str_detect(.data$from,"^BASE"))
# Make an network object using `asNetwork` to convert the igraph class into
# network class. Then `ggnetwork` to the universal dataframe format used in
# `ggplot2`
set.seed(1)
network.graph <- igraph::graph_from_data_frame(
      network.data.edges,
      directed = TRUE,
      vertices = network.data.vertices) %>%
      intergraph::asNetwork() %>% 
      ggnetwork(layout = "fruchtermanreingold")

# An finally plot using 

ggplot(network.graph, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges( color = "grey50") +
  geom_nodes(aes(color = tag.type)) +
  theme_blank()


```


## Network Plots
```{r MakeNetworkPlot}
# Reload the file to make sure that there is no unintended contamination from
# earlier tests.

test.xml.file.path <- system.file("extdata","Example1.xml", package = "MBSE.utils")
test.xml <- xml2::read_xml(test.xml.file.path)

# Find triples with specific encoding
find.triple <- c("Requirement1",
                 "Function",
                 "System",
                 "System Element 1",
                 "System Element 2",
                 "InterfaceBlock1",
                 "Exchange Item 1",
                 "Exchange Item 2",
                 "Sub Function A",
                 "Sub Function B",
                 "block",
                 "requirement",
                 "satisfy",
                 "refine",
                 "allocate",
                 "input",
                 "output",
                 "incoming",
                 "model" ,
                 "outgoing",
                 "ProxyPort" ,
                 "links" )

# my.test.triple <-  triples.out %>%   filter(datumValue %in% find.triple)
search.ontology  <-   knitr::combine_words(find.triple)

graph.cols.to.include <- c("id", "datumEntity","tag.type", 
                           "nominal.name", "name", 
                           "plot.core","plot.size")
 
```

```{r LoadExample1XML }

# We need to find the path to the directory, not the file.
dir.path <- fs::path_dir(test.xml.file.path)

xml.info <- xml.file.to.triples(dir.path, file.number = 1, 
                                xpath = "/*") # original /* "//uml:Model"

triples.out <- xml.info$triples
this.file.name <-  paste(xml.info$file.name, "xpath = /*" )
pander::pandoc.header(this.file.name, level = 2)
```

```{r Example1Xml, eval=TRUE, fig.cap= "A network representation of Example1.xml, with XML structure dropped.", fig.width=7,fig.height=5}
triples.plus <- add.ontology.col(triples = triples.out,
                                       triples.to.find = find.triple,
                            datumAttributeCol = "plot.core",
                            datumValueCol = "Ontology term",
                            datumValueNot = "XML tag"
                                 )

triples.plus <- add.ontology.col(triples = triples.plus,
                                       triples.to.find = find.triple,
                            datumAttributeCol = "plot.size",
                            datumValueCol = "5",
                            datumValueNot = "1"
                                 )




# Make the triples into a network
network.graph <- triples.plus %>% triples.to.context.network(drop.xml.edges = TRUE, cols.to.include = graph.cols.to.include) 
 # %>% save.graph.ml(  name.prefix = "rootXML")
  
# Now convert the network graph to a dataframe
 network.graph.df <-  network.graph %>%
      intergraph::asNetwork() %>% 
      ggnetwork::ggnetwork(layout = "fruchtermanreingold") %>% tidyr::drop_na()
 
  network.graph.df$plot.size <- network.graph.df$plot.size %>% as.numeric()

ggplot2::ggplot(network.graph.df, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges( color = "grey50") +
    geom_nodes(aes(colour = plot.core, size = plot.size )) +
  geom_text(aes(label= nominal.name), data = network.graph.df[network.graph.df$plot.core %in% "Ontology term", ]) +

  theme_blank(legend.position="bottom")



```



